Global Climate and Atmospheric Composition of the Ultra-Hot Jupiter WASP-103\lowercase{b} from \HST\ and \Spitzer\ Phase Curve Observations

### 2018-05-02 ApJ_submit_v3/ms.tex
second submission

WASP-103b is an archetype of the class of ultra-hot Jupiters, with orbital periods of about one day and dayside temperatures typically $>2000\,\mathrm{K}$. These very hot planets were initially predicted to have inverted temperature pressure profiles due to strong optical absorption by TiO/VO in the upper atmospheres \citep{hubeny03, fortney08}; however, observations of their emission spectra have been inconclusive on their thermal structure and composition. In the near-infrared, where water is the dominant absorber, some spectra show water absorption features, some show emission features, and some are consistent with blackbody models \citep{madhusudhan11, crossfield12, stevenson14b, haynes15, evans16, beatty17a, beatty17b, sheppard17, arcangeli18, mansfield18}.  A variety of explanations have been proposed for these results, including low metallicity or high carbon-to-oxygen compositions, dayside clouds, and finely tuned isothermal temperature pressure profiles. Recently, \cite{arcangeli18} showed that water dissociation and H- opacity on the hot dayside play an important role in the atmospheres of these ultra-hot planets and may be responsible for some of the blackbody-like near-IR spectra. In this work, we put these results in context by investigating the global thermal structure and composition of the ultra-hot Jupiter WASP-103b.

@@

To maximize the possibility of detecting WASP-103's rotation, we normalized the photometry such that each observing season has the same mean, thereby removing any long-term variability in WASP-103 and/or the comparison stars (Figure\,\ref{fig:photometry}, top panel).  To estimate the stellar rotation period, we performed a periodogram analysis of the normalized data set based on least-squares fitting of sine curves.  The resulting frequency spectrum and the phase curve computed with the best period are shown in the middle and lower panels of Figure\,\ref{fig:photometry} respectively. The best-fit period is 6.814 days, which agrees closely with the estimated stellar rotation period of 6.855 days \citep[based on the projected stellar rotation velocity and stellar radius reported in][]{gillon14}. The peak-to-peak variability amplitude is 0.005 mag.  Based on the formalism in \cite{zellem17}, we calculate that this variability will bias the measured eclipse depth by $\lesssim10$\,parts per million (ppm) from epoch to epoch, which is well below the photon-limited precision of our measurements.

@@

@@

 \newcommand{\noop}[1]{}
\begin{thebibliography}{}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi
\providecommand{\url}[1]{\href{#1}{#1}}

\bibitem[{Adcroft {et~al.}(2004)Adcroft, Campin, Hill, \&
  Marshall}]{Adcroft2004}
Adcroft, A., Campin, J.-M., Hill, C., \& Marshall, J. 2004, Monthly Weather
  Review, 132, 2845.
\newblock \url{https://doi.org/10.1175/MWR2823.1}

@@

\bibitem[{Braak(2006)}]{braak06}
Braak, C. J. F.~T. 2006, Statistics and Computing, 16, 239.
\newblock \url{http://dx.doi.org/10.1007/s11222-006-8769-1}

@@

\bibitem[{Jones {et~al.}(2001)Jones, Oliphant, Peterson, \&
  Others}]{jones_scipy_2001}
Jones, E., Oliphant, T., Peterson, P., \& Others. 2001, {SciPy}: Open source
  scientific tools for Python, , .
\newblock \url{http://www.scipy.org/}

\bibitem[{Kass \& Raftery(1995)}]{kass95}
Kass, R.~E., \& Raftery, A.~E. 1995, Journal of the American Statistical
  Association, 90, 773.
\newblock
  \url{http://amstat.tandfonline.com/doi/abs/10.1080/01621459.1995.10476572}

@@

\bibitem[{{Kreidberg} {et~al.}(2014{\natexlab{a}}){Kreidberg}, {Bean},
  {D{\'e}sert}, {Benneke}, {Deming}, {Stevenson}, {Seager}, {Berta-Thompson},
  {Seifahrt}, \& {Homeier}}]{kreidberg14a}
{Kreidberg}, L., {Bean}, J.~L., {D{\'e}sert}, J.-M., {et~al.}
  2014{\natexlab{a}}, \nat, 505, 69

\bibitem[{{Kreidberg} {et~al.}(2014{\natexlab{b}}){Kreidberg}, {Bean},
  {D{\'e}sert}, {Line}, {Fortney}, {Madhusudhan}, {Stevenson}, {Showman},
  {Charbonneau}, {McCullough}, {Seager}, {Burrows}, {Henry}, {Williamson},
  {Kataria}, \& {Homeier}}]{kreidberg14b}
---. 2014{\natexlab{b}}, \apjl, 793, L27

@@

@@

\bibitem[{{Mansfield} {et~al.}(\noop{3001}submitted){Mansfield}, {Bean},
  {Line}, {Paramentier}, \& {Kreidberg}}]{mansfield18}
{Mansfield}, M., {Bean}, J.~B., {Line}, M., {Paramentier}, V., \& {Kreidberg},
  L. \noop{3001}submitted, Submitted to \apj

@@

\bibitem[{{Parmentier} {et~al.}(\noop{3001}submitted){Parmentier}, {Line},
  {Bean}, {Fortney}, {D\'{e}sert}, \& {Kreidberg}}]{parmentier18}
{Parmentier}, V., {Line}, M.~R., {Bean}, J., {et~al.} \noop{3001}submitted,
  Submitted to A\&A

>>>>

WASP-103b is an archetype of the class of ultra-hot Jupiters, with orbital
periods of about one day and dayside temperatures typically $>2000\,\mathrm{K}$.
These very hot planets were initially predicted to have inverted temperature
pressure profiles due to strong optical absorption by TiO/VO in the upper
atmospheres \citep{hubeny03, fortney08}; however, observations of their emission
spectra have been inconclusive on their thermal structure and composition. In
the near-infrared, where water is the dominant absorber, some spectra show water
absorption features, some show emission features, and some are consistent with
blackbody models \citep{madhusudhan11, crossfield12, stevenson14b, haynes15,
evans16, beatty17a, beatty17b, sheppard17, arcangeli18, mansfield18}.  A variety
of explanations have been proposed for these results, including low metallicity
or high carbon-to-oxygen compositions, dayside clouds, and finely tuned
isothermal temperature pressure profiles. Recently, \cite{arcangeli18,
lothringer18} showed that water dissociation and H- opacity on the hot dayside play an important role in the atmospheres of these ultra-hot planets and may be responsible for some of the blackbody-like near-IR spectra. In this work, we put these results in context by investigating the global thermal structure and composition of the ultra-hot Jupiter WASP-103b.

@@

To maximize the possibility of detecting WASP-103's rotation, we normalized the
photometry such that each observing season has the same mean, thereby removing
any long-term variability in WASP-103 and/or the comparison stars
(Figure\,\ref{fig:photometry}, top panel).  To estimate the stellar rotation
period, we performed a periodogram analysis of the normalized data set based on
least-squares fitting of sine curves.  The resulting frequency spectrum and the
phase curve computed with the best period are shown in the middle and lower
panels of Figure\,\ref{fig:photometry} respectively. The best-fit period is
6.814 days, which agrees closely with the estimated stellar rotation period of
6.855 days \citep[based on the projected stellar rotation velocity and stellar
radius reported in][]{gillon14}. \added{There are two nearby peaks in the
periodogram (panel b of Figure\,\ref{fig:photometry}) that are one-year aliases
of each other, and we chose the peak that better matches the stellar rotation velocity.} The peak-to-peak variability amplitude is 0.005 mag.  Based on the formalism in \cite{zellem17}, we calculate that this variability will bias the measured eclipse depth by $\lesssim10$\,parts per million (ppm) from epoch to epoch, which is well below the photon-limited precision of our measurements.

@@

\added{As a test, we also calculated the difference in brightness temperature
between the \HST\ and \Spitzer\ 4.5 $\mu$m data (reported in
Table\,\ref{table:teffs}) for both nightside phases ($\phi = 0.1$ and 0.9). At
phase 0.1 and 0.9, the \Spitzer\ temperature is lower by 2.7 and 5.6\,$\sigma$,
respectively.  These values are a lower limit to the significance, because the
brightness temperatures noted in the table also include the uncertainty in the
stellar $T_\mathrm{eff}$ (which increases the uncertainty on the absolute planet
temperature but not the relative temperatures that are relevant for this
calculation). The drop in brightness temperature is more significant at phase 0.9 than it is at 0.1, providing further evidence in addition to the phase 0.1 retrieval that the nightside temperature pressure profile is not inverted.}

@@

 \newcommand{\noop}[1]{}
\begin{thebibliography}{}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi

\bibitem[{Adcroft {et~al.}(2004)Adcroft, Campin, Hill, \&
  Marshall}]{Adcroft2004}
Adcroft, A., Campin, J.-M., Hill, C., \& Marshall, J. 2004, Monthly Weather
  Review, 132, 2845

@@

\bibitem[{Braak(2006)}]{braak06}
Braak, C. J. F.~T. 2006, Statistics and Computing, 16, 239

@@

\bibitem[{Jones {et~al.}(2001)Jones, Oliphant, Peterson, \&
  Others}]{jones_scipy_2001}
Jones, E., Oliphant, T., Peterson, P., \& Others. 2001, {SciPy}: Open source
  scientific tools for Python, ,

\bibitem[{Kass \& Raftery(1995)}]{kass95}
Kass, R.~E., \& Raftery, A.~E. 1995, Journal of the American Statistical
  Association, 90, 773

@@

\bibitem[{{Kreidberg} {et~al.}(2014{\natexlab{a}}){Kreidberg}, {Bean},
  {D{\'e}sert}, {Line}, {Fortney}, {Madhusudhan}, {Stevenson}, {Showman},
  {Charbonneau}, {McCullough}, {Seager}, {Burrows}, {Henry}, {Williamson},
  {Kataria}, \& {Homeier}}]{kreidberg14b}
{Kreidberg}, L., {Bean}, J.~L., {D{\'e}sert}, J.-M., {et~al.}
  2014{\natexlab{a}}, \apjl, 793, L27

\bibitem[{{Kreidberg} {et~al.}(2014{\natexlab{b}}){Kreidberg}, {Bean},
  {D{\'e}sert}, {Benneke}, {Deming}, {Stevenson}, {Seager}, {Berta-Thompson},
  {Seifahrt}, \& {Homeier}}]{kreidberg14a}
---. 2014{\natexlab{b}}, \nat, 505, 69

@@

\bibitem[{{Lothringer} {et~al.}(2018){Lothringer}, {Barman}, \&
  {Koskinen}}]{lothringer18}
{Lothringer}, J.~D., {Barman}, T., \& {Koskinen}, T. 2018, ArXiv e-prints,
  arXiv:1805.00038

@@

\bibitem[{{Mansfield} {et~al.}(2018){Mansfield}, {Bean}, {Line}, {Parmentier},
  {Kreidberg}, {Desert}, {Fortney}, {Stevenson}, {Arcangeli}, \&
  {Dragomir}}]{mansfield18}
{Mansfield}, M., {Bean}, J.~L., {Line}, M.~R., {et~al.} 2018, ArXiv e-prints,
  arXiv:1805.00424

@@

\bibitem[{{Parmentier} {et~al.}(2018){Parmentier}, {Line}, {Bean}, {Mansfield},
  {Kreidberg}, {Lupu}, {Visscher}, {Desert}, {Fortney}, {Deleuil}, {Arcangeli},
  {Showman}, \& {Marley}}]{parmentier18}
{Parmentier}, V., {Line}, M.~R., {Bean}, J.~L., {et~al.} 2018, ArXiv e-prints,
  arXiv:1805.00096
